A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-06-04 LensingPaper_draft_v4.01.tex
Corrected typo.
Done editing section 3.3.1 on the theory behind JAM.

%_____FORMAT_____
\usepackage{hyperref} %ready-to-lick referencing

@@

%--------------------------------------------
\Wilma{[Referee says: "The manuscript in its current form is also slightly longer than necessary at certain
points: [...] section
3.3.1 contains a lot of material that is implicit in Cappellari's JAM code, and
therefore does not need to be explained equation by equation here."]}
\OLD{JAM modelling assumes galaxies to be (i) collisionless (i.e., the collisionless Boltzmann equation for the distribution function $f(\vect{x},\vect{v},t)$ has to be satisfied, $\frac{\diff f(\vect{x},\vect{v},t)}{\diff t} = 0$), (ii) in a steady state ($\frac{\partial}{\partial t} = 0$), (iii) axisymmetric (best described in cylindrical coordinates $(R,z,\phi)$ and $\frac{\partial}{\partial \phi} = 0$). From this follow the axisymmetric Jeans equations as the vector-valued first moment of the Boltzmann equation, i.e.,
\begin{equation*}
\int \vect{v} \frac{\diff f}{\diff t} \Diff3 v = 0.
\end{equation*}}
\NEW{JAM modelling assumes galaxies to be (i) collisionless, (ii) in a steady state and (iii) axisymmetric. With these assumptions the axisymmetric Jeans equations follow from the vector-valued first moment of the collisionless Boltzmann equation (see Equations (4.221)-(4.222) in \Wilma{???}). The Jeans equations are functions of the second velocity moments $\langle v_i v_j\rangle$ (with $i,j\in(R,z,\phi)$), the stellar tracer number density \Wilma{Continue????????}} To be able to solve the Jeans equations, additional assumptions about the velocity ellipsoid tensor $\langle v_i v_j\rangle$ have to be made. We follow \citet{Cap08} and assume firstly that the galaxy's velocity ellipsoid is aligned with the cylindrical coordinate system, i.e., $\langle v_R v_z\rangle = \langle v_z v_\phi \rangle = \langle v_\phi v_R \rangle = 0$. Secondly, we assume a constant ratio between the radial and vertical second velocity moments, 
\begin{equation}
\beta_z \equiv 1 - \langle v_z^2 \rangle / \langle v_R^2\rangle, \label{eq:bz}
\end{equation}
where $\beta_z$ is the velocity anisotropy parameter. This reduces the Jeans equations to two equations for $\langle v_z^2 \rangle$ and $\langle v_\phi^2 \rangle$ which can be solved by means of one integration,
\begin{eqnarray}
n \langle v_z^2 \rangle (R,z) &=& \int_0^\infty n \frac{\partial \Phi}{\partial z} \diff z \label{eq:Jeans1}\\
n \langle v_\phi^2 \rangle (R,z) &=& R \frac{\partial}{\partial R} \left( \frac{n \langle v_z^2 \rangle}{1-\beta_z} \right) + \frac{n \langle v_z^2 \rangle}{1 - \beta_z} + R n \frac{\partial \Phi}{\partial R},\label{eq:Jeans2}
\end{eqnarray}
where $n(\vect{x}) \equiv \int f(\vect{x},\vect{v}) \Diff3 v$ is the number density of (stellar) tracers and $\Phi(\vect{x})$ the galaxy's gravitational potential generated by the mass density $\rho(\vect{x})$.

The JAM modelling approach by \citet{Cap08} expresses the tracer and mass density in terms of MGEs (see also \citealt{1994A&A...285..723E}). The tracer density $n(\vect{x})$ is assumed to be proportional to the observed and deprojected brightness distribution $\nu(R,z)$ in Equation \eqref{eq:deprojMGE}. The mass density $\rho(R,z)$ can consist of several sets of MGEs, describing stellar and DM components. The MGE for $\Phi(R,z)$ is generated from the mass density MGE by integrating the Poisson equation \citep{1994A&A...285..723E}. Equations \eqref{eq:Jeans1} and \eqref{eq:Jeans2} together with \eqref{eq:bz} provide the velocity dispersion tensor $\langle v_i v_j \rangle(R,z) = \langle v_i^2 \rangle(R,z)$ (with $i,j \in \{ R,\phi, z\}$ and $\langle v_i v_j \rangle=0$ if $i\neq j$). $\langle v_i^2 \rangle$ is then rotated by the inclination angle $i$ to the coordinate system of the observer; $(x',y')$ is the plane of the sky and $z'$ the line-of-sight, where $x'$ is alinged with the galaxy's major axis. Taking a light-weighted projection along the line-of-sight gives a model prediction for the line-of-sight velocity second moment, which is comparable to actual spectroscopic measurements of the second velocity moment. Details of the derivation using the MGE formalism are given in \citet{Cap08} and in the appendix of \citet{GlennEC}. We therefore just give the result for the line-of-sight second velocity moment prediction from the Jeans equations,
\begin{eqnarray}
&&\left(I \langle v_\text{los}^2\rangle\right)(x',y')\nonumber\\
&&= 4 \pi^{3/2} G \int_0^1 \sum_{i=1}^N \sum_{k=1}^M \nu_{0,i} \ q_k \ \rho_{0,k} \ u^2\nonumber\\
&& \times \frac{\sigma_i^2 \ q^{2}_i \left( \cos^2 i + \frac{\sin^2 i}{1-\beta_{z,i}}\right) + \mathscr{D} \  {x'}^{2} \sin^2 i}{(1-\mathscr{C}u^2) \sqrt{(\mathscr{A} + \mathscr{B} \cos^2 i) \left[1-(1-q^{2}_k) u^2 \right]}}\nonumber\\
&& \times \exp\left\{- \mathscr{A} \left[{x'}^{2} + \frac{(\mathscr{A}+\mathscr{B}) \ {y'}^{2}}{\mathscr{A}+\mathscr{B}\cos^2 i}\right] \right\} \diff u, \label{eq:explicitLOSvelCap}
\end{eqnarray}
with $N$ Gaussians describing the tracer distribution $\nu(R,z)$ and $M$ Gaussians describing the mass distribution $\rho(R,z)$, $\rho_{0,k}$ being the $k$-th mass density Gaussian evaluated at $(R=0,z=0)$ and
\begin{eqnarray*}
\mathscr{A} &=& \frac 12 \left(\frac{u^2}{\sigma_k^2} + \frac{1}{\sigma_i^2} \right)\nonumber\\
\mathscr{B} &=& \frac 12 \left\{\frac{1-q^{2}_i}{\sigma_i^2 q^{2}_i} + \frac{(1-q^{2}_k)u^4}{\sigma_k^2 \left[1-(1-{q}_k^{2})u^2 \right]} \right\}\nonumber\\
\mathscr{C} &=& 1- q^{2}_k - \frac{\sigma_i^2 q^{2}_i}{\sigma^2_k}\nonumber\\
\mathscr{D} &=& 1 - \frac{q^{2}_i}{1-\beta_{z,i}} - \left[ \left(1-\frac{1}{1-\beta_{i,z}}\right)\mathscr{C} + \frac{1-q^{2}_k}{1-\beta_{z,i}}\right] u^2.\nonumber
\end{eqnarray*}
The JAM modelling code by \citet{Cap08} evaluates Equation \eqref{eq:explicitLOSvelCap} numerically for a given luminous tracer and mass distribution MGE and a given inclination.
%--------------------------------------------

>>>>

%_____FORMAT_____
\usepackage{hyperref} %ready-to-click referencing

@@

%--------------------------------------------
\OLD{JAM modelling assumes galaxies to be (i) collisionless (i.e., the collisionless Boltzmann equation for the distribution function $f(\vect{x},\vect{v},t)$ has to be satisfied, $\frac{\diff f(\vect{x},\vect{v},t)}{\diff t} = 0$), (ii) in a steady state ($\frac{\partial}{\partial t} = 0$), (iii) axisymmetric (best described in cylindrical coordinates $(R,z,\phi)$ and $\frac{\partial}{\partial \phi} = 0$). From this follow the axisymmetric Jeans equations as the vector-valued first moment of the Boltzmann equation, i.e.,
\begin{equation*}
\int \vect{v} \frac{\diff f}{\diff t} \Diff3 v = 0.
\end{equation*}}
\NEW{JAM modelling assumes galaxies to be (i) collisionless, (ii) in a steady state and (iii) axisymmetric. With these assumptions the axisymmetric Jeans equations follow from the vector-valued first moment of the collisionless Boltzmann equation (see Equations (4.221)-(4.222) in \citet{2008gady.book.....B}). The Jeans equations are functions of the second velocity moments $\langle v_i v_j\rangle$ (with $i,j\in(R,z,\phi)$), the number density of (stellar) tracers $n(\vect{x})$, and the galaxy's gravitational potential $\Phi(\vect{x})$ generated by the mass density $\rho(\vect{x})$.}

To be able to solve the Jeans equations, additional assumptions about the velocity ellipsoid tensor $\langle v_i v_j\rangle$ have to be made. We follow \citet{Cap08} and assume firstly that the galaxy's velocity ellipsoid is aligned with the cylindrical coordinate system, i.e., $\langle v_R v_z\rangle = \langle v_z v_\phi \rangle = \langle v_\phi v_R \rangle = 0$. Secondly, we assume a constant ratio between the radial and vertical second velocity moments, 
\begin{equation}
\beta_z \equiv 1 - \langle v_z^2 \rangle / \langle v_R^2\rangle, \label{eq:bz}
\end{equation}
where $\beta_z$ is the velocity anisotropy parameter. \OLD{This reduces the Jeans equations to two equations for $\langle v_z^2 \rangle$ and $\langle v_\phi^2 \rangle$ which can be solved by means of one integration,
\begin{eqnarray}
n \langle v_z^2 \rangle (R,z) &=& \int_0^\infty n \frac{\partial \Phi}{\partial z} \diff z \label{eq:Jeans1}\\
n \langle v_\phi^2 \rangle (R,z) &=& R \frac{\partial}{\partial R} \left( \frac{n \langle v_z^2 \rangle}{1-\beta_z} \right) + \frac{n \langle v_z^2 \rangle}{1 - \beta_z} + R n \frac{\partial \Phi}{\partial R},\label{eq:Jeans2}
\end{eqnarray}
where $n(\vect{x}) \equiv \int f(\vect{x},\vect{v}) \Diff3 v$ is the number density of (stellar) tracers and $\Phi(\vect{x})$ the galaxy's gravitational potential generated by the mass density $\rho(\vect{x})$.}

The JAM modelling approach by \citet{Cap08} expresses the tracer and mass density in terms of MGEs (see also \citealt{1994A&A...285..723E}). The tracer density $n(\vect{x})$ is assumed to be proportional to the observed and deprojected brightness distribution $\nu(R,z)$ in Equation \eqref{eq:deprojMGE}. The mass density $\rho(R,z)$ can consist of several sets of MGEs, describing stellar and DM components. The MGE for $\Phi(R,z)$ is generated from the mass density MGE by integrating the Poisson equation \citep{1994A&A...285..723E}. \OLD{Equations \eqref{eq:Jeans1} and \eqref{eq:Jeans2} together with \eqref{eq:bz} provide the velocity dispersion tensor $\langle v_i v_j \rangle(R,z) = \langle v_i^2 \rangle(R,z)$ (with $i,j \in \{ R,\phi, z\}$ and $\langle v_i v_j \rangle=0$ if $i\neq j$).}

\NEW{In this way one can now calculate an unambiguous model prediction for the velocity dispersion tensor $\langle v_i v_j \rangle(R,z) = \langle v_i^2 \rangle(R,z)$ (with $i,j \in \{ R,\phi, z\}$ and $\langle v_i v_j \rangle=0$ if $i\neq j$). To compare it with observations, $\langle v_i^2 \rangle(R,z)$ has to be}\OLD{$\langle v_i^2 \rangle$ is then} rotated by the inclination angle $i$ to the coordinate system of the observer; $(x',y')$ is the plane of the sky and $z'$ the line-of-sight, where $x'$ is alinged with the galaxy's major axis. Taking a light-weighted projection along the line-of-sight gives a model prediction for the line-of-sight velocity second moment \NEW{$\langle v_\text{los}^2\rangle(x',y')$}, which is comparable to actual spectroscopic measurements of the second velocity moment. Details of the derivation \NEW{from the Jeans equations} using the MGE formalism are \OLD{given}\NEW{laid out} in \citet{Cap08} and in the appendix of \citet{GlennEC}. \OLD{We therefore just give the result for the line-of-sight second velocity moment prediction from the Jeans equations,}\NEW{The result for $\langle v_\text{los}^2\rangle(x',y')$ in particular is given in Equation (28) and Equation (A18), respectively.}
\OLD{\begin{eqnarray}
&&\left(I \langle v_\text{los}^2\rangle\right)(x',y')\nonumber\\
&&= 4 \pi^{3/2} G \int_0^1 \sum_{i=1}^N \sum_{k=1}^M \nu_{0,i} \ q_k \ \rho_{0,k} \ u^2\nonumber\\
&& \times \frac{\sigma_i^2 \ q^{2}_i \left( \cos^2 i + \frac{\sin^2 i}{1-\beta_{z,i}}\right) + \mathscr{D} \  {x'}^{2} \sin^2 i}{(1-\mathscr{C}u^2) \sqrt{(\mathscr{A} + \mathscr{B} \cos^2 i) \left[1-(1-q^{2}_k) u^2 \right]}}\nonumber\\
&& \times \exp\left\{- \mathscr{A} \left[{x'}^{2} + \frac{(\mathscr{A}+\mathscr{B}) \ {y'}^{2}}{\mathscr{A}+\mathscr{B}\cos^2 i}\right] \right\} \diff u, \label{eq:explicitLOSvelCap}
\end{eqnarray}
with $N$ Gaussians describing the tracer distribution $\nu(R,z)$ and $M$ Gaussians describing the mass distribution $\rho(R,z)$, $\rho_{0,k}$ being the $k$-th mass density Gaussian evaluated at $(R=0,z=0)$ and
\begin{eqnarray*}
\mathscr{A} &=& \frac 12 \left(\frac{u^2}{\sigma_k^2} + \frac{1}{\sigma_i^2} \right)\nonumber\\
\mathscr{B} &=& \frac 12 \left\{\frac{1-q^{2}_i}{\sigma_i^2 q^{2}_i} + \frac{(1-q^{2}_k)u^4}{\sigma_k^2 \left[1-(1-{q}_k^{2})u^2 \right]} \right\}\nonumber\\
\mathscr{C} &=& 1- q^{2}_k - \frac{\sigma_i^2 q^{2}_i}{\sigma^2_k}\nonumber\\
\mathscr{D} &=& 1 - \frac{q^{2}_i}{1-\beta_{z,i}} - \left[ \left(1-\frac{1}{1-\beta_{i,z}}\right)\mathscr{C} + \frac{1-q^{2}_k}{1-\beta_{z,i}}\right] u^2.\nonumber
\end{eqnarray*}}

The JAM modelling code by \citet{Cap08} evaluates \OLD{Equation \eqref{eq:explicitLOSvelCap}}\NEW{this expression for $\langle v_\text{los}^2\rangle(x',y')$} numerically for a given luminous tracer and mass distribution MGE and a given inclination.
\Wilma{[TO DO: MAke sure M and N are introduced somewhere.]}
%--------------------------------------------
